---
title: "Demography of M.tb Lineages"
author: "Mary O'Neill"
date: "November 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(ape)
#source("https://bioconductor.org/biocLite.R")
#biocLite("ggtree")
require("ggtree")
require(phytools)
require(phangorn)
require(cowplot)
require(rworldmap)
require(reshape2)
require(RColorBrewer)
require(data.table)
require(splitstackshape)
library(viridis)
library(colorRamps)
library(scales)
require(akima)
require(dplyr)
require(ggplot2)

opts_knit$set(root.dir = "C://Users/Mary/PepLab/data/Phylogeography/Phylo_Figures_Data/")         
setwd("C://Users/Mary/PepLab/data/Phylogeography/Phylo_Figures_Data/")
```

## Comparison of BSP from country and UN lineage specific analyses

The lineage specific country analyses were done back in February from individual lineage SNP alignments with the rate fixed to 1. The lineage specific UN region analyses were done in November from the global SNP alignment, fixing the rate to 1 but also accounting for invariant sites. To explore whether this effected the BSP results I will plot them together.

```{r bsp, echo=FALSE, warning=FALSE}
l1.co.bsp <- read.table("lin1_BSP.data", skip=1, header=T)
l2.co.bsp <- read.table("lin2_BSP.data", skip=1, header=T)
l3.co.bsp <- read.table("lin3_BSP.data", skip=1, header=T)
l4.co.bsp <- read.table("lin4_BSP.data", skip=1, header=T)

l1.un.bsp <- read.table("lin1_UN_BSP.data", skip=1, header=T)
l2.un.bsp <- read.table("lin2_UN_BSP.data", skip=1, header=T)
l3.un.bsp <- read.table("lin3_UN_BSP.data", skip=1, header=T)
l4.un.bsp <- read.table("lin4_UN_BSP.data", skip=1, header=T)
l5.un.bsp <- read.table("lin5_UN_BSP.data", skip=1, header=T)
l6.un.bsp <- read.table("lin6_UN_BSP.data", skip=1, header=T)
l7.un.bsp <- read.table("lin7_UN_BSP.data", skip=1, header=T)


#######Lineage 1

#Scale the BSP's to reflect a MRCA for lin1 of 2353
X1 = 2353/max(l1.co.bsp$Time)
#l1.co.bsp$TimeScaled <- l1.co.bsp$Time * X
l1.co.scaled <- l1.co.bsp * X1 / 1.3

Y1 = 2352/max(l1.un.bsp$Time)
#l1.un.bsp$TimeScaled <- l1.un.bsp$Time * Y
l1.un.scaled <- l1.un.bsp * Y1 / 1.3


#Combine the BSPs for lineage 1
l1.bsp <- rbind(data.frame(analysis="co", l1.co.scaled), 
                data.frame(analysis="un", l1.un.scaled))

#Plot
l1.bsp.p <- ggplot(l1.bsp) + geom_line(aes(Time, Mean, group=analysis, colour=analysis)) + geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower, group=analysis, colour=analysis, fill=analysis), alpha=.2) + scale_color_manual(values = c("gray", "blue")) + scale_fill_manual(values = c("gray", "blue")) + xlab("Years Before Present") + ylab("Eff Pop Size") + scale_x_reverse(limits = c(2000, 0)) + scale_y_log10(limits=c(10,1e6)) + geom_hline(yintercept=7571.8, linetype="dashed", color = "red") + ggtitle("Lineage 1")


#######Lineage 2

#Scale the BSP's to reflect a MRCA for lin2 of 2021
X2 = 2021/max(l2.co.bsp$Time)
#l2.co.bsp$TimeScaled <- l2.co.bsp$Time * X
l2.co.scaled <- l2.co.bsp * X2 / 1.3

Y2 = 2021/max(l2.un.bsp$Time)
#l2.un.bsp$TimeScaled <- l2.un.bsp$Time * Y
l2.un.scaled <- l2.un.bsp * Y2 / 1.3


#Combine the BSPs for lineage 1
l2.bsp <- rbind(data.frame(analysis="co", l2.co.scaled), 
                data.frame(analysis="un", l2.un.scaled))

#Plot
l2.bsp.p <- ggplot(l2.bsp) + geom_line(aes(Time, Mean, group=analysis, colour=analysis)) + geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower, group=analysis, colour=analysis, fill=analysis), alpha=.2) + scale_color_manual(values = c("gray", "blue")) + scale_fill_manual(values = c("gray", "blue")) + xlab("Years Before Present") + ylab("Eff Pop Size") + scale_x_reverse(limits = c(2000, 0)) + scale_y_log10(limits=c(10,1e6)) + geom_hline(yintercept=4485.66, linetype="dashed", color = "red") + ggtitle("Lineage 2")


#######Lineage 3

#Scale the BSP's to reflect a MRCA for lin3 of 1482
X3 = 1482/max(l3.co.bsp$Time)
#l3.co.bsp$TimeScaled <- l3.co.bsp$Time * X
l3.co.scaled <- l3.co.bsp * X3 / 1.3

Y3 = 1482/max(l3.un.bsp$Time)
#l3.un.bsp$TimeScaled <- l3.un.bsp$Time * Y
l3.un.scaled <- l3.un.bsp * Y3 / 1.3


#Combine the BSPs for lineage 1
l3.bsp <- rbind(data.frame(analysis="co", l3.co.scaled), 
                data.frame(analysis="un", l3.un.scaled))

#Plot
l3.bsp.p <- ggplot(l3.bsp) + geom_line(aes(Time, Mean, group=analysis, colour=analysis)) + geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower, group=analysis, colour=analysis, fill=analysis), alpha=.2) + scale_color_manual(values = c("gray", "blue")) + scale_fill_manual(values = c("gray", "blue")) + xlab("Years Before Present") + ylab("Eff Pop Size") + scale_x_reverse(limits = c(2000, 0)) + scale_y_log10(limits=c(10,1e6)) + geom_hline(yintercept=3987.45, linetype="dashed", color = "red") + ggtitle("Lineage 3")


#######Lineage 4

#Scale the BSP's to reflect a MRCA for lin4 of 1918
X4 = 1918/max(l4.co.bsp$Time)
#l4.co.bsp$TimeScaled <- l4.co.bsp$Time * X
l4.co.scaled <- l4.co.bsp * X4 / 1.3

Y4 = 1918/max(l4.un.bsp$Time)
#l4.un.bsp$TimeScaled <- l4.un.bsp$Time * Y
l4.un.scaled <- l4.un.bsp * Y4 / 1.3


#Combine the BSPs for lineage 1
l4.bsp <- rbind(data.frame(analysis="co", l4.co.scaled), 
                data.frame(analysis="un", l4.un.scaled))

#Plot
l4.bsp.p <- ggplot(l4.bsp) + geom_line(aes(Time, Mean, group=analysis, colour=analysis)) + geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower, group=analysis, colour=analysis, fill=analysis), alpha=.2) + scale_color_manual(values = c("gray", "blue")) + scale_fill_manual(values = c("gray", "blue")) + xlab("Years Before Present") + ylab("Eff Pop Size") + scale_x_reverse(limits = c(2000, 0)) + scale_y_log10(limits=c(10,1e6)) + geom_hline(yintercept=7797.08, linetype="dashed", color = "red") + ggtitle("Lineage 4")


plot_grid(
  l1.bsp.p,
  l2.bsp.p,
  l3.bsp.p,
  l4.bsp.p,
  ncol=2, align = "h"
)


#######Lineage 5

#Scale the BSP's to reflect a MRCA for lin5 of 1211

Y5 = 1211/max(l5.un.bsp$Time)
#l5.un.bsp$TimeScaled <- l5.un.bsp$Time * Y
l5.un.scaled <- l5.un.bsp * Y4 / 1.3


#Combine the BSPs for lineage 1
l5.bsp <- rbind(data.frame(analysis="un", l5.un.scaled))

#Plot
l5.bsp.p <- ggplot(l5.bsp) + geom_line(aes(Time, Mean, group=analysis, colour=analysis)) + geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower, group=analysis, colour=analysis, fill=analysis), alpha=.2) + scale_color_manual(values = c("blue")) + scale_fill_manual(values = c("blue")) + xlab("Years Before Present") + ylab("Eff Pop Size") + scale_x_reverse(limits = c(2000, 0)) + scale_y_log10(limits=c(10,1e6)) + geom_hline(yintercept=7797.08, linetype="dashed", color = "red") + ggtitle("Lineage 5")

#######Lineage 6

#Scale the BSP's to reflect a MRCA for lin5 of 1900

Y6 = 1900/max(l6.un.bsp$Time)
#l6.un.bsp$TimeScaled <- l6.un.bsp$Time * Y
l6.un.scaled <- l6.un.bsp * Y4 / 1.3


#Combine the BSPs for lineage 1
l6.bsp <- rbind(data.frame(analysis="un", l6.un.scaled))

#Plot
l6.bsp.p <- ggplot(l6.bsp) + geom_line(aes(Time, Mean, group=analysis, colour=analysis)) + geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower, group=analysis, colour=analysis, fill=analysis), alpha=.2) + scale_color_manual(values = c("blue")) + scale_fill_manual(values = c("blue")) + xlab("Years Before Present") + ylab("Eff Pop Size") + scale_x_reverse(limits = c(2000, 0)) + scale_y_log10(limits=c(10,1e6)) + geom_hline(yintercept=7797.08, linetype="dashed", color = "red") + ggtitle("Lineage 6")

#######Lineage 7

#Scale the BSP's to reflect a MRCA for lin7 of 687

Y7 = 687/max(l7.un.bsp$Time)
#l7.un.bsp$TimeScaled <- l7.un.bsp$Time * Y
l7.un.scaled <- l7.un.bsp * Y4 / 1.3


#Combine the BSPs for lineage 1
l7.bsp <- rbind(data.frame(analysis="un", l7.un.scaled))

#Plot
l7.bsp.p <- ggplot(l7.bsp) + geom_line(aes(Time, Mean, group=analysis, colour=analysis)) + geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower, group=analysis, colour=analysis, fill=analysis), alpha=.2) + scale_color_manual(values = c("blue")) + scale_fill_manual(values = c("blue")) + xlab("Years Before Present") + ylab("Eff Pop Size") + scale_x_reverse(limits = c(2000, 0)) + scale_y_log10(limits=c(10,1e6)) + geom_hline(yintercept=7797.08, linetype="dashed", color = "red") + ggtitle("Lineage 7")


plot_grid(
  l1.bsp.p,
  l2.bsp.p,
  l3.bsp.p,
  l4.bsp.p,
  l5.bsp.p,
  l6.bsp.p,
  l7.bsp.p,
  ncol=2, align = "h"
)

```

The BSPs generated in the two analyses are in strong agreement. I fit the time scales to relect the timings from the (corrected) TMRCA timings from the original full analysis. Note that the two analyses for which ESS values were low (L2, L4) are still running with longer chain lengths.

###Dadi Results

```{r dadi-L1, echo=FALSE, warning=FALSE}

##Lineage 1
l1.exp.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin1/likelihood_grid_expansion.txt", header=TRUE)
l1.exp <-interp2xyz(interp(x=l1.exp.pre$nu, y=l1.exp.pre$T, z=l1.exp.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l1.exp) <- c("vB", "t", "LL")
l1.exp <- tbl_df(l1.exp)
l1.exp <- filter(l1.exp, !is.na(LL))

l1.gro.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin1/likelihood_grid_growth.txt", header=TRUE)
l1.gro <-interp2xyz(interp(x=l1.gro.pre$nu, y=l1.gro.pre$T, z=l1.gro.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l1.gro) <- c("vB", "t", "LL")
l1.gro <- tbl_df(l1.gro)
l1.gro <- filter(l1.gro, !is.na(LL))

l1.lim <- c(min(l1.exp$LL), max(l1.exp$LL), min(l1.gro$LL), max(l1.gro$LL))

l1.exp.p <- ggplot(l1.exp, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 300, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l1.lim), max(l1.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("L1 - expansion") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none") +
  geom_point(aes(x=66.92842, y=0.8031205), colour="red")


l1.gro.p <- ggplot(l1.gro, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 300, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l1.lim), max(l1.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("L1 - growth") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none") +
  geom_point(aes(x=286.8294, y=2.030231), colour="red")

plot_grid(
  l1.exp.p,
  l1.gro.p,
  ncol=2, align = "h"
)
```

```{r dadi-L2, echo=FALSE, warning=FALSE}

##Lineage 2
l2.exp.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin2/likelihood_grid_expansion.txt", header=TRUE)
l2.exp <-interp2xyz(interp(x=l2.exp.pre$nu, y=l2.exp.pre$T, z=l2.exp.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l2.exp) <- c("vB", "t", "LL")
l2.exp <- tbl_df(l2.exp)
l2.exp <- filter(l2.exp, !is.na(LL))

l2.gro.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin2/likelihood_grid_growth.txt", header=TRUE)
l2.gro <-interp2xyz(interp(x=l2.gro.pre$nu, y=l2.gro.pre$T, z=l2.gro.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l2.gro) <- c("vB", "t", "LL")
l2.gro <- tbl_df(l2.gro)
l2.gro <- filter(l2.gro, !is.na(LL))

l2.lim <- c(min(l2.exp$LL), max(l2.exp$LL), min(l2.gro$LL), max(l2.gro$LL))

l2.exp.p <- ggplot(l2.exp, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 300, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l2.lim), max(l2.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l2 - expansion") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none")+
  geom_point(aes(x=145.7558, y=0.2745687), colour="red")


l2.gro.p <- ggplot(l2.gro, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 300, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l2.lim), max(l2.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l2 - growth") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none")+
  geom_point(aes(x=538.6626, y=1.245329), colour="red")

plot_grid(
  l2.exp.p,
  l2.gro.p,
  ncol=2, align = "h"
)


```

```{r dadi-L3, echo=FALSE, warning=FALSE}

##Lineage 3
l3.exp.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin3/likelihood_grid_expansion.txt", header=TRUE)
l3.exp <-interp2xyz(interp(x=l3.exp.pre$nu, y=l3.exp.pre$T, z=l3.exp.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l3.exp) <- c("vB", "t", "LL")
l3.exp <- tbl_df(l3.exp)
l3.exp <- filter(l3.exp, !is.na(LL))

l3.gro.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin3/likelihood_grid_growth.txt", header=TRUE)
l3.gro <-interp2xyz(interp(x=l3.gro.pre$nu, y=l3.gro.pre$T, z=l3.gro.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l3.gro) <- c("vB", "t", "LL")
l3.gro <- tbl_df(l3.gro)
l3.gro <- filter(l3.gro, !is.na(LL))

l3.lim <- c(min(l3.exp$LL), max(l3.exp$LL), min(l3.gro$LL), max(l3.gro$LL))

l3.exp.p <- ggplot(l3.exp, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 300, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l3.lim), max(l3.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l3 - expansion") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none")+
  geom_point(aes(x=104.157, y=2.499924), colour="red")


l3.gro.p <- ggplot(l3.gro, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 20, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l3.lim), max(l3.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l3 - growth") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none")+
  geom_point(aes(x=8499.866, y=86.36667), colour="red")

plot_grid(
  l3.exp.p,
  l3.gro.p,
  ncol=2, align = "h"
)
```

```{r dadi-L4, echo=FALSE, warning=FALSE}

##Lineage 4
l4.exp.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin4/likelihood_grid_expansion.txt", header=TRUE)
l4.exp <-interp2xyz(interp(x=l4.exp.pre$nu, y=l4.exp.pre$T, z=l4.exp.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l4.exp) <- c("vB", "t", "LL")
l4.exp <- tbl_df(l4.exp)
l4.exp <- filter(l4.exp, !is.na(LL))

l4.gro.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin4/likelihood_grid_growth.txt", header=TRUE)
l4.gro <-interp2xyz(interp(x=l4.gro.pre$nu, y=l4.gro.pre$T, z=l4.gro.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l4.gro) <- c("vB", "t", "LL")
l4.gro <- tbl_df(l4.gro)
l4.gro <- filter(l4.gro, !is.na(LL))

l4.lim <- c(min(l4.exp$LL), max(l4.exp$LL), min(l4.gro$LL), max(l4.gro$LL))

l4.exp.p <- ggplot(l4.exp, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 300, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l4.lim), max(l4.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l4 - expansion") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none")+
  geom_point(aes(x=58.56261, y=0.6366894), colour="red")


l4.gro.p <- ggplot(l4.gro, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 1500, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l4.lim), max(l4.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l4 - growth") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none")+
  geom_point(aes(x=335.9563, y=3.554986), colour="red")

plot_grid(
  l4.exp.p,
  l4.gro.p,
  ncol=2, align = "h"
)
```

```{r dadi-L5, echo=FALSE, warning=FALSE}

##Lineage 5
l5.exp.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin5/likelihood_grid_expansion.txt", header=TRUE)
l5.exp <-interp2xyz(interp(x=l5.exp.pre$nu, y=l5.exp.pre$T, z=l5.exp.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l5.exp) <- c("vB", "t", "LL")
l5.exp <- tbl_df(l5.exp)
l5.exp <- filter(l5.exp, !is.na(LL))

l5.gro.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin5/likelihood_grid_growth.txt", header=TRUE)
l5.gro <-interp2xyz(interp(x=l5.gro.pre$nu, y=l5.gro.pre$T, z=l5.gro.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l5.gro) <- c("vB", "t", "LL")
l5.gro <- tbl_df(l5.gro)
l5.gro <- filter(l5.gro, !is.na(LL))

l5.lim <- c(min(l5.exp$LL), max(l5.exp$LL), min(l5.gro$LL), max(l5.gro$LL))

l5.exp.p <- ggplot(l5.exp, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 300, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l5.lim), max(l5.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l5 - expansion") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none")+
  geom_point(aes(x=443.6372, y=4.937224), colour="red")


l5.gro.p <- ggplot(l5.gro, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 10, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l5.lim), max(l5.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l5 - growth") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none")+
  geom_point(aes(x=9862.081, y=7.8161), colour="red")

plot_grid(
  l5.exp.p,
  l5.gro.p,
  ncol=2, align = "h"
)
```

```{r dadi-L6, echo=FALSE, warning=FALSE}

##Lineage 6
l6.exp.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin6/likelihood_grid_expansion.txt", header=TRUE)
l6.exp <-interp2xyz(interp(x=l6.exp.pre$nu, y=l6.exp.pre$T, z=l6.exp.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l6.exp) <- c("vB", "t", "LL")
l6.exp <- tbl_df(l6.exp)
l6.exp <- filter(l6.exp, !is.na(LL))

l6.gro.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin6/likelihood_grid_growth.txt", header=TRUE)
l6.gro <-interp2xyz(interp(x=l6.gro.pre$nu, y=l6.gro.pre$T, z=l6.gro.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l6.gro) <- c("vB", "t", "LL")
l6.gro <- tbl_df(l6.gro)
l6.gro <- filter(l6.gro, !is.na(LL))

l6.lim <- c(min(l6.exp$LL), max(l6.exp$LL), min(l6.gro$LL), max(l6.gro$LL))

l6.exp.p <- ggplot(l6.exp, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 300, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l6.lim), max(l6.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l6 - expansion") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none")+
  geom_point(aes(x=57.51868, y=0.9785816), colour="red")


l6.gro.p <- ggplot(l6.gro, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 300, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l6.lim), max(l6.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l6 - growth") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none")+
  geom_point(aes(x=288.9614, y=1.361122), colour="red")

plot_grid(
  l6.exp.p,
  l6.gro.p,
  ncol=2, align = "h"
)
```

```{r dadi-L7, echo=FALSE, warning=FALSE}

##Lineage 7
l7.exp.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin7/likelihood_grid_expansion.txt", header=TRUE)
l7.exp <-interp2xyz(interp(x=l7.exp.pre$nu, y=l7.exp.pre$T, z=l7.exp.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l7.exp) <- c("vB", "t", "LL")
l7.exp <- tbl_df(l7.exp)
l7.exp <- filter(l7.exp, !is.na(LL))

l7.gro.pre <- read.table("C://Users/Mary/PepLab/data/Phylogeography/dadi/upperbound_1000_100/lin7/likelihood_grid_growth.txt", header=TRUE)
l7.gro <-interp2xyz(interp(x=l7.gro.pre$nu, y=l7.gro.pre$T, z=l7.gro.pre$LL, duplicate="mean"), data.frame=TRUE)
colnames(l7.gro) <- c("vB", "t", "LL")
l7.gro <- tbl_df(l7.gro)
l7.gro <- filter(l7.gro, !is.na(LL))

l7.lim <- c(min(l7.exp$LL), max(l7.exp$LL), min(l7.gro$LL), max(l7.gro$LL))

l7.exp.p <- ggplot(l7.exp, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 300, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l7.lim), max(l7.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l7 - expansion") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none")+
  geom_point(aes(x=18.93599, y=3.269871), colour="red")


l7.gro.p <- ggplot(l7.gro, aes(x = vB, y = t, z = LL, fill = LL)) + 
  geom_tile() + 
  stat_contour(colour = "black", alpha=0.25, bins = 500, na.rm=TRUE) + 
  scale_fill_viridis(option = "inferno", limits = c(min(l7.lim), max(l7.lim))) +
  #scale_y_continuous(limits=c(0, 10)) +
  #scale_x_continuous(limits=c(0, 200)) +
  theme_bw() +
  ggtitle("l7 - growth") +
  theme(text = element_text(size=20),
        legend.text = element_text(size=18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        panel.background=element_blank(),
        panel.grid.major= element_blank(),
        legend.position="none") +
  geom_point(aes(x=368.665, y=91.60298), colour="red")

plot_grid(
  l7.exp.p,
  l7.gro.p,
  ncol=2, align = "h"
)
```

###Tree Figure

```{r trees, echo=FALSE, warning=FALSE}

l1t <- read.beast("lin1_UN_MCC.tree")
l1 <- fortify(l1t)
l1.dat <- l1[,c('node', 'parent', 'label', 'geography', 'geography.prob', 'geography.set', 'geography.set.prob', 'posterior', 'height_median')]

l1.dat$geography.set.prob = gsub("\\{|\\}", "", l1.dat$geography.set.prob)
l1.dat$geography.set = gsub("\\{|\\}", "", l1.dat$geography.set)

#subset the tips
tre_tips <- subset(l1, isTip == TRUE, c(label, geography))

#split the tips by UN
tre_iso = split(tre_tips$label, tre_tips$geography)

tre.g <- groupOTU(l1t, tre_iso)

ggtree(tre.g, layout="unrooted") + geom_tippoint(aes(color=group)) 
  #geom_text2(aes(subset = isTip, label = geography, color = group), hjust=-0.1)  
  

l2t <- read.beast("lin2_UN_MCC.tree")
l2 <- fortify(l2t)
l2.dat <- l2[,c('node', 'parent', 'label', 'geography', 'geography.prob', 'geography.set', 'geography.set.prob', 'posterior', 'height_median')]

l2.dat$geography.set.prob = gsub("\\{|\\}", "", l2.dat$geography.set.prob)
l2.dat$geography.set = gsub("\\{|\\}", "", l2.dat$geography.set)

#subset the tips
tre_tips <- subset(l2, isTip == TRUE, c(label, geography))

#split the tips by UN
tre_iso = split(tre_tips$label, tre_tips$geography)

tre.g <- groupOTU(l2t, tre_iso)

ggtree(tre.g, layout="unrooted") + geom_tippoint(aes(color=group)) 
  #geom_text2(aes(subset = isTip, label = geography, color = group), hjust=-0.1)  

```


```{r mig_bsp}

l1.un.bsp <- read.table("lin1_UN_BSP.data", skip=1, header=T)
l2.un.bsp <- read.table("lin2_UN_BSP.data", skip=1, header=T)
l3.un.bsp <- read.table("lin3_UN_BSP.data", skip=1, header=T)
l4.un.bsp <- read.table("lin4_UN_BSP.data", skip=1, header=T)



l1t <- read.beast("lin4_UN_MCC.tree")

l1 <- fortify(l1t)

l1.dat <- l1[,c('node', 'parent', 'label', 'geography', 'geography.prob', 'geography.set', 'geography.set.prob', 'posterior', 'height_median')]

ggtree(l1t)

#lin1
d1 <- read.table("161111_lin1_UNMig.txt", header=F, sep="\t", na.strings=c("na","NA"))
names(d1) <- c('node','parent','node_un','node_prob','parent_un','parent_prob','mig','node_med','parent_med','node_posterior','parent_posterior')  


#scale factor for lineage 1: 133534.8


Y1 = 2352/max(d1$node_med, na.rm=TRUE)
Z1 = 45576/max(d1$node_med, na.rm=TRUE)

d1mod <- d1[!is.na(d1[,'node_med']),]
d1mod <- d1mod[!is.na(d1mod[,'parent_med']),]
d1mod$node_med <- d1mod$node_med * Y1
d1mod$parent_med <- d1mod$parent_med * Y1



rr <- range(d1mod[,c('node_med', 'parent_med')])
bb <- seq(floor(rr[1]),ceiling(rr[2]))
mm <- apply(d1mod[,c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))


cl1 <- apply(d1mod[,c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))
l1 <- apply(cl1,1,sum)
cl1un <- apply(d1mod[d1mod$mig == "Yes",c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))
l1un <- apply(cl1un,1,sum)


lin1 <- cbind(data.frame(bb), data.frame(l1), data.frame(l1un))
lin1.m <- melt(lin1, id.vars="bb")


l1.mig.p <- ggplot(lin1[lin1$l1 > 4,]) + geom_line(aes(bb, l1un/l1), colour = "black") + theme(legend.position = "none") + xlab("") + ylab("") + scale_x_reverse() + geom_ribbon(aes(x=bb, ymax=((l1un/l1) + 1/l1), ymin=((l1un/l1) - 1/l1)), alpha=.2) + scale_fill_manual(values=c("gray")) + scale_y_continuous(limits=c(-.2,.7))

+ geom_ribbon(aes(x=bb, ymax=((l2un/l2) + 1/l2), ymin=((l2un/l2) - 1/l2)), alpha=.2) + scale_fill_manual(values=c("gray"))

l1.un.scaled <- l1.un.bsp * Y1

l1.bsp.p <- ggplot(l1.un.scaled) + geom_line(aes(Time, Mean)) + geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower), alpha=.2) + scale_color_manual(values = c("gray")) + scale_fill_manual(values = c("gray")) + xlab("") + ylab("") + scale_x_reverse() + scale_y_continuous() + geom_hline(yintercept=7571.8, linetype="dashed", color = "red") + ylab("") + xlab("")


plot_grid(
  l1.mig.p, #+ xlim(1910, 0),
  l1.bsp.p, #+ xlim(1910, 0),
  ncol=1,
  align="v"
)








#lin2
d2 <- read.table("161111_lin2_UNMig.txt", header=F, sep="\t", na.strings=c("na","NA"))
names(d2) <- c('node','parent','node_un','node_prob','parent_un','parent_prob','mig','node_med','parent_med','node_posterior','parent_posterior')  


Y2 = 2021/max(d2$node_med, na.rm=TRUE)
Z2 = 39173/max(d2$node_med, na.rm=TRUE)

d2mod <- d2[!is.na(d2[,'node_med']),]
d2mod <- d2mod[!is.na(d2mod[,'parent_med']),]
d2mod$node_med <- d2mod$node_med * Z2
d2mod$parent_med <- d2mod$parent_med * Z2



rr <- range(d2mod[,c('node_med', 'parent_med')])
bb <- seq(floor(rr[1]),ceiling(rr[2]))
mm <- apply(d2mod[,c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))


cl2 <- apply(d2mod[,c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))
l2 <- apply(cl2,1,sum)
cl2un <- apply(d2mod[d2mod$mig == "Yes",c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))
l2un <- apply(cl2un,1,sum)

lin2 <- cbind(data.frame(bb), data.frame(l2), data.frame(l2un))
lin2.m <- melt(lin2, id.vars="bb")



l2.mig.p <- ggplot(lin2[lin2$l2 > 4,]) + geom_line(aes(bb, l2un/l2), colour = "black") + theme(legend.position = "none") + xlab("") + ylab("") + scale_x_reverse() + geom_ribbon(aes(x=bb, ymax=((l2un/l2) + 1/l2), ymin=((l2un/l2) - 1/l2)), alpha=.2) + scale_fill_manual(values=c("gray")) + scale_y_continuous(limits = c(-0.2, 0.7))



l2.un.scaled <- l2.un.bsp * Z2

l2.bsp.p <- ggplot(l2.un.scaled) + geom_line(aes(Time, Mean)) + geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower), alpha=.2) + scale_color_manual(values = c("gray")) + scale_fill_manual(values = c("gray")) + xlab("") + ylab("") + scale_x_reverse() + scale_y_continuous() + geom_hline(yintercept=4486, linetype="dashed", color = "red") + ylab("") + xlab("")


b <- plot_grid(
  l2.mig.p + xlim(1100, 0),
  l2.bsp.p + xlim(1100, 0),
  ncol=1,
  align="v"
)


#lin3
d3 <- read.table("161111_lin3_UNMig.txt", header=F, sep="\t", na.strings=c("na","NA"))
names(d3) <- c('node','parent','node_un','node_prob','parent_un','parent_prob','mig','node_med','parent_med','node_posterior','parent_posterior')  


Y3 = 1482/max(d3$node_med, na.rm=TRUE)
Z3 = 28722/max(d3$node_med, na.rm=TRUE)

d3mod <- d3[!is.na(d3[,'node_med']),]
d3mod <- d3mod[!is.na(d3mod[,'parent_med']),]
d3mod$node_med <- d3mod$node_med * Z3
d3mod$parent_med <- d3mod$parent_med * Z3



rr <- range(d3mod[,c('node_med', 'parent_med')])
bb <- seq(floor(rr[1]),ceiling(rr[2]))
mm <- apply(d3mod[,c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))


cl3 <- apply(d3mod[,c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))
l3 <- apply(cl3,1,sum)
cl3un <- apply(d3mod[d3mod$mig == "Yes",c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))
l3un <- apply(cl3un,1,sum)

lin3 <- cbind(data.frame(bb), data.frame(l3), data.frame(l3un))
lin3.m <- melt(lin3, id.vars="bb")


l3.mig.p <- ggplot(lin3[lin3$l3 > 4,]) + geom_line(aes(bb, l3un/l3), colour = "black") + theme(legend.position = "none") + xlab("") + ylab("") + scale_x_reverse() + geom_ribbon(aes(x=bb, ymax=((l3un/l3) + 1/l3), ymin=((l3un/l3) - 1/l3)), alpha=.2) + scale_fill_manual(values=c("gray")) + scale_y_continuous(limits = c(-0.2, 0.7))



l3.un.scaled <- l3.un.bsp * Z3

l3.bsp.p <- ggplot(l3.un.scaled) + geom_line(aes(Time, Mean)) + geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower), alpha=.2) + scale_color_manual(values = c("gray")) + scale_fill_manual(values = c("gray")) + xlab("") + ylab("") + scale_x_reverse() + scale_y_continuous() + geom_hline(yintercept=3987, linetype="dashed", color = "red") + ylab("") + xlab("")


c <- plot_grid(
  l3.mig.p + xlim(1300, 0),
  l3.bsp.p + xlim(1300, 0),
  ncol=1,
  align="v"
)








#lin 4
d4 <- read.table("161111_lin4_UNMig.txt", header=F, sep="\t", na.strings=c("na","NA"))
names(d4) <- c('node','parent','node_un','node_prob','parent_un','parent_prob','mig','node_med','parent_med','node_posterior','parent_posterior')  


Y4 = 1918/max(d4$node_med, na.rm=TRUE)
Z4 = 37161/max(d4$node_med, na.rm=TRUE)

d4mod <- d4[!is.na(d4[,'node_med']),]
d4mod <- d4mod[!is.na(d4mod[,'parent_med']),]
d4mod$node_med <- d4mod$node_med * Z4
d4mod$parent_med <- d4mod$parent_med * Z4



rr <- range(d4mod[,c('node_med', 'parent_med')])
bb <- seq(floor(rr[1]),ceiling(rr[2]))
mm <- apply(d4mod[,c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))


cl4 <- apply(d4mod[,c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))
l4 <- apply(cl4,1,sum)
cl4un <- apply(d4mod[d4mod$mig == "Yes",c('parent_med', 'node_med')],1,function(x) bb %in% seq(floor(x["node_med"]),ceiling(x["parent_med"])))
l4un <- apply(cl4un,1,sum)

lin4 <- cbind(data.frame(bb), data.frame(l4), data.frame(l4un))
lin4.m <- melt(lin4, id.vars="bb")


l4.mig.p <- ggplot(lin4[lin4$l4 > 4,]) + geom_line(aes(bb, l4un/l4), colour = "black") + theme(legend.position = "none") + xlab("") + ylab("") + scale_x_reverse() + geom_ribbon(aes(x=bb, ymax=((l4un/l4) + 1/l4), ymin=((l4un/l4) - 1/l4)), alpha=.2) + scale_fill_manual(values=c("gray")) + scale_y_continuous(limits = c(-0.2, 0.7))



l4.un.scaled <- l4.un.bsp * Z4

l4.bsp.p <- ggplot(l4.un.scaled) + geom_line(aes(Time, Mean)) + geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower), alpha=.2) + scale_color_manual(values = c("gray")) + scale_fill_manual(values = c("gray")) + xlab("") + ylab("") + scale_x_reverse() + scale_y_log10(limits=c(10,1e6)) + geom_hline(yintercept=7797, linetype="dashed", color = "red") + ylab("") + xlab("")

l4.bsp.2 <- ggplot(l4.un.scaled) + 
  geom_line(aes(Time, Mean)) + 
  geom_ribbon(aes(x=Time, ymax=Upper, ymin=Lower), alpha=.2) + 
  scale_color_manual(values = c("gray")) + 
  scale_fill_manual(values = c("gray")) + xlab("Years Before Present") + 
  ylab("Eff Pop Size") + scale_x_reverse() + scale_y_continuous() + 
  geom_hline(yintercept=7797, linetype="dashed", color = "red") + ylab("") + xlab("")

d <- plot_grid(
  l4.mig.p + xlim(1600, 0),
  #l4.bsp.p + xlim(2500, 0),
  l4.bsp.2 + xlim(1600, 0),
  ncol=1,
  align="v"
)

fig <- plot_grid(
  l1.mig.p + xlim(1910, 0),
  l2.mig.p + xlim(1100, 0),
  l1.bsp.p + xlim(1910, 0),
  l2.bsp.p + xlim(1100, 0),
  l3.mig.p + xlim(1300, 0),
  l4.mig.p + xlim(1600, 0),
  l3.bsp.p + xlim(1300, 0),
  l4.bsp.2 + xlim(1600, 0),
  ncol=2,
  align="v"
)

fig2 <- plot_grid(
  l1.mig.p + xlim(37015.5, 0),
  l2.mig.p + xlim(21317.83, 0),
  l1.bsp.p + xlim(37015.5, 0),
  l2.bsp.p + xlim(21317.83, 0),
  l3.mig.p + xlim(25193.8, 0),
  l4.mig.p + xlim(31007.75, 0),
  l3.bsp.p + xlim(25193.8, 0),
  l4.bsp.2 + xlim(31007.75, 0),
  ncol=2,
  align="v"
)


exportPlot(fig2, "migration_bsp2", 9, 9)


```

